Modelling biological pathway dynamics\\with \tas

### 2013-07-10 Paper.tex
Added grant number
Two small changes in Q&A. One in the paper

We acknowledge support from Netherlands Organisation for Scientific Research (NWO) CASIMIR to J. Scholma.}%
}

@@

It has been observed that the activation of extracellular regulated kinase (ERK) shows significantly different 
dynamics upon treatment with EGF as compared to treatment with NGF. A transient, peak-shaped 
activation is observed when PC-12 cells are treated with EGF, whereas treatment with NGF results in sustained activation of ERK,
even after removal of the input signals via growth factor-neutralizing antibodies. These activation profiles are tied to different cellular 
outcomes, with transient activation leading to proliferation and sustained activation leading to differentiation. 
These findings led the authors of~\cite{egf-ngf} to hypothesize the existence of a positive feedback from ERK to an upstream node in the network and 
existence of a new network component that blocks this feedback. It was found that NGF prevents this blocking, leading to sustained activation~\cite{egf-ngf}.
We have formalized and implemented these topological reasonings in an ANIMO model, shown in Figure~\ref{fig:case-study-model}. 
The parameters used in this network are given in Table~\ref{tab:case-study-parameters}.
The accompanying experimental conditions are the addition of either EGF or NGF
in sufficient quantity to fully activate their respective receptors. After 10 minutes a growth factor-neutralizing 
antibody is added, shutting off the input signal to the network. The evolution of the network is
observed for 60 minutes from the initial treatment.

>>>>

We acknowledge support from Netherlands Organisation for Scientific Research (NWO) Casimir grant 018.003.031 to J. Scholma.}%
}

@@

It has been observed that the activation of extracellular regulated kinase (ERK) shows significantly different 
dynamics upon treatment with EGF as compared to treatment with NGF. A transient, peak-shaped 
activation is observed when PC-12 cells are treated with EGF, whereas treatment with NGF results in sustained activation of ERK,
even after removal of the input signals via growth factor-neutralizing antibodies. These activation profiles are tied to different cellular 
outcomes, with transient activation leading to proliferation and sustained activation leading to differentiation. 
These findings led the authors of~\cite{egf-ngf} to hypothesize the existence of a positive feedback from ERK to an upstream node in the network and 
existence of a new network component that blocks this feedback. It was found that NGF prevents this blocking, leading to sustained activation~\cite{egf-ngf}.
We have formalized and implemented these topological reasonings in an ANIMO model, shown in Figure~\ref{fig:case-study-model}. 
The parameters used in this network are given in Table~\ref{tab:case-study-parameters}.
The corresponding experimental conditions are the addition of either EGF or NGF
in sufficient quantity to fully activate their respective receptors. After 10 minutes a growth factor-neutralizing 
antibody is added, shutting off the input signal to the network. The evolution of the network is
observed for 60 minutes from the initial treatment.
